### 2019-10-11 introduction.tex
intro Aditya comments addressed. ref added for Google sheet

%!TEX root = main.tex

%\IEEEraisesectionheading{\section{Introduction}}
\section{Introduction}
\label{sec:intro}
\new{With a user base of more than  one-tenth of the world's population,
{\bf \em spreadsheets are by far the most popular medium for ad-hoc exploration
and analysis of data}~\cite{excel-users}.} 
Studies show that information workers prefer to operate on their data within spreadsheets while shunning enterprise solutions 
with more advanced analytical features~\cite{chan1996use,raden2005shedding}. 
One popular joke among those developing business intelligence applications is that  the ``export to excel'' button is the third-most commonly used button from the menu bar, after OK and Cancel~\cite{bakke2016expressive}.
Spreadsheets enable users to view, structure, and present data
in an intuitive tabular layout, wherein users can map their data and tasks;
this tabular layout is essential to
the popularity of spreadsheets~\cite{nardi1990spreadsheet}.

Using this tabular layout effectively involves
navigation, \ie
{\em ``the process of viewing and manipulating the computer display to
show another portion of the information space''}~\cite{nav-study}.
Navigation is supported
via two unit operations, scrolling and steering.
Scrolling is the action of moving displayed text or graphics up, down, or across a computer screen, in order to view different parts of the spreadsheet.
For example, when analyzing data, users may scroll
to compare data across different screens, or to get
a high-level view of the overall spreadsheet.
Steering, on the other hand, involves clicking the left mouse button and then dragging the mouse pointer through the spreadsheet to select a specific region. For example, to issue a formula,
users may steer to select
the subset of the data
to be operated on as an argument within the formula.
Most frequently used spreadsheet formulae
require users to perform steering actions~\cite{bradbard2014spreadsheet, lawson2009comparison}.
Overall, both scrolling and steering are crucial as
users navigate spreadsheets to
identify, compare, and summarize data.

\begin{figure*}
     \centering %trim=left bottom right top
  %\includegraphics[width=\linewidth,trim={0 95 0 80},clip]{images/navigation.pdf}
  \includegraphics[width=\linewidth,trim={0 95 0 80},clip]{images/screenshot.pdf}
    \caption{\noah: navigation interface consisting of (a) a zoomable overview and (b) an aggregate column integrated with (c) a spreadsheet. A context bar consisting of (d) a navigation history displaying locations visited so far using the overview, and (e) a breadcrumb showing the current navigation path (\eg Home). (f) The user’s current focus in the spreadsheet is highlighted on the overview. (g) Columns corresponding to the navigation attribute (city) and aggregate column (availability) are highlighted on the spreadsheet.}
  \label{fig:ux}
\end{figure*}

\new{
However, navigating spreadsheets using scrolling or steering
is challenging, since spreadsheet data span multiple screens,
making it hard to synthesize, analyze, makes sense of, or
operate on it~\cite{network-context,nardi1990spreadsheet}.
With the ease of data generation, and with spreadsheets
now supporting increasingly larger datasets, \eg Google Sheets
now supports five million cells~\cite{gs}, a $12.5X$ increase
from the previous limit of 400K cells, 
{\em navigating
data within spreadsheets is only becoming even harder},
thanks to
multiple inter-related reasons: 
}
\squishlist
\item {\em Loss of overview and context.}
When navigating
spreadsheets, users can easily lose the context
of where they are and where they should go next~\cite{network-context}.
The only navigational context provided by spreadsheets
is the built-in scrollbar that acts as a one-dimensional
overview and indicates the user's current location on the sheet.
However, since this overview does not capture the
layout and structure of the data,
users are forced to mentally assimilate
the layout and recall it on-demand, as they navigate a spreadsheet.

\item {\em Cognitive and mechanical burdens.}
The lack of contextual cues leads to severe
cognitive and mechanical burdens
for users~\cite{cockburn2009review}.
Users often end up taking their own drastic measures to
avoid getting lost; for example,
some users create personalized overviews
extrinsic to the spreadsheet,
by sketching maps
of spreadsheets on paper~\cite{network-context}.
Other users add their own landmarks such as headers or
colored cells, as a visual affordance to assist
in navigation~\cite{network-context}.
Steering via dragging the mouse pointer across multiple screens to select
a subset of data as input to a formula
can often be challenging as well:
the only remedy is for users to abandon steering entirely and
instead remember the range of
the subset of data of interest, and then correctly enter this range
as the argument to the formula, often giving rise to errors that are increasingly prevalent in spreadsheets~\cite{panko1998we}.

\item {\em Visual discontinuities.}
The limited viewport afforded to the user
introduces a visual discontinuity between the
information being displayed.
For example, comparing spatially separated subsets
of data within the spreadsheet
requires moving back and forth between multiple viewports,
which can be overwhelming~\cite{nardi1990spreadsheet,network-context}.
As an alternative, users tend to copy subsets of data side by side to
reduce the visual discontinuity~\cite{nardi1990spreadsheet,network-context},
which is cumbersome.
\squishend

\noindent
Overall,  while navigating present-day spreadsheets, 
{\bf \em users often lose context,
get overwhelmed, and experience visual discontinuities.} 
Addressing these challenges requires considerable manual effort.
As we \new{will} argue in Section~\ref{sec:related},
existing spreadsheet features such as
pivot tables, named ranges, and subtotals,
\new{partially alleviate 
some of the aforementioned 
challenges but do not eliminate them entirely.
For example, pivot tables generate a summary
while losing the correspondence between the
raw data and the summary, while named ranges require users
to manually associate names with ranges of data.}

So, how do we support more effective navigation of data
within spreadsheets?
One approach would be to try to integrate
an overview of the overall structure of the data along
with the spreadsheet~\cite{grudin2001partitioning}
\new{resulting in} a classical
overview+detail interface
where the spreadsheet is the detailed view.
Overview+detail interfaces are used
to facilitate navigation in various domains such as 
text editors and maps~\cite{cockburn2009review}.
\new{Users can manipulate
the overview or detailed view,
to perform high-level or low-level operations,
respectively.}
Overview+detail interfaces
have been shown to be effective in these domains,
reducing cognitive load for users by
providing \new{them} the big picture first,
helping them quickly assimilate 
the information space~\cite{cockburn2009review}.
\new{Our goal is to {\em integrate an overview plug-in 
with spreadsheets} that captures
the overall structure of the data,
while supporting interactions that address
the difficulties in typical navigational 
operations like scrolling and steering.
{\bf \em It is essential
that our interface is a plug-in that 
enhances the capabilities of spreadsheets
that so many users
are used to and reliant on,
as opposed to a potentially jarring or confusing 
replacement for spreadsheets.}
}
% \saj{In this paper, our goal is to integrate an overview plug-in with spreadsheets that captures the overall structure of the data while allowing interactions that minimize the difficulties caused by typical navigational operations like scrolling and steering.}

However, while an overview plug-in for spreadsheets
does seem appealing and natural,
developing it leads to several challenges.

\squishlist
\item {\em Overview modality.}
One approach for an overview is to simply add a
zoomed out version of the entire sheet as a pane on the side,
as in popular presentation software like 
Microsoft PowerPoint or text editors
like Sublime Text.
\newsaj{The zoomed out overview
displays the entire data 
at a lower magnification scale.} 
Unfortunately, this approach
would not suffice for a spreadsheet.
\new{An overview should provide a
comprehensible big picture view:
for a \newsaj{spreadsheet of numbers, text, or formulae}, 
when zooming out beyond a point, 
\newsaj{an overview displayed at 
such low magnification scale} 
would be unreadable.}
Another approach, adopted by map tools like the
early versions of Google maps,
is to use the overview
to provide a global context of the user's current location
being displayed in the zoomed in detailed view~\cite{cockburn2009review}.
While the overview remains static, users can perform semantic zooming operations~\cite{perlin1993pad} on the detailed
view which allows objects to be represented differently at different scales. Since spreadsheets already display the raw data,
zooming into and out of a detailed view consisting of
this raw data is not meaningful. 
\new{How do we design an overview to dynamically 
change as users 
seek a more fine-grained or coarse-grained view 
of the overall structure of the data?}

\item {\em Construction of the overview.}
Given a spreadsheet
with many rows, 
one approach to constructing a dynamic overview 
is mapping rows of data to high level groups, 
similar to online maps. 
In online maps, cities are grouped into states 
and states are grouped into countries, 
forming a multi-granularity hierarchy. 
How do we automatically group spreadsheet 
rows together in a similar ``meaningful''
way
such that this grouping applies to all 
data types, including strings and numbers? 
If the automatically generated grouping is 
not semantically meaningful, how do we allow the users
to customize the grouping modality? 
How do we facilitate interactions that enable 
users to view the overview at multiple granularities?

\item {\em Operations on the overview.}
Following the construction of a dynamic overview, 
the next challenge is to design simple interactions 
that achieve similar outcomes as scrolling and steering.
For example, an alternative to scrolling 
can be to leverage the groups of the overview 
to access the rows mapped to that group. 
\newsaj{As the granularity of the dynamic overview 
changes, how do we efficiently update the mapping 
from spreadsheet rows to the 
finer or coarser groups 
so that scrolling remains seamless? 
Similarly, how do we leverage
the overview to steer spreadsheet data (i.e., select a range of data)
for formula computation? How do we present the results of
the formula within the overview? One approach can be to adopt
the pivot table-like presentation of results. Within their summary view, 
pivot tables display 
aggregate formula results (\eg \code{SUM}, \code{COUNT}) 
alongside each group. 
However, unlike pivot tables, users can explore the overview at
multiple-granularities. As the granularity changes,
the grouping of rows also changes; 
making the previous formula results 
inconsistent with the new groups.
How do we 
recompute the results of a formula 
as the granularity changes 
without requiring the users to 
reissue the formula from scratch?}
\agp{the last three questions are fairly confusing and assume that you already know what NOAH does.. can we simplify?}

\item {\em Seamless integration as a plug-in.}
\newsaj{Finally, how can we 
design an overview with a generic set of features,
that can be integrated
with any existing spreadsheet tool on any 
dataset without impacting
existing functionalities or look-and-feel?
How do we ensure that the interactions 
supported by the overview 
are 
consistent with
traditional spreadsheet semantics,
and complement existing spreadsheet interactions?
How to we enable coordinated interactions across both views, \ie 
the overview and the raw spreadsheet, such that they
remain consistent at any given time?}
\squishend
\agp{A general comment: can we ensure that either we use
questions for all of the items above or none of them for uniformity?}

\stitle{\noah: \new{a navigation plug-in for spreadsheets}.} We address the aforementioned challenges
in \noah,
an in-situ
\textbf{\underline{n}}avigation interface
for \textbf{\underline{o}}verviewing
and \textbf{\underline{a}}nalyzing
spreadsheet data \textbf{\underline{h}}olistically.
\noah is constructed as a plugin to
an existing spreadsheet tool,
{\scshape DataSpread}~\cite{dataspread},
an open-source scalable web-based spreadsheet.
While \noah's design is not tied to {\sc DataSpread},
we opted not to use other popular spreadsheet
tools like Google Sheets and Microsoft Excel because they
are closed source.
Figure~\ref{fig:ux} shows a snapshot of \noah.
When the user chooses to explore
the data by a specific attribute,
a multi-granularity overview is constructed and displayed
within \noah, next to the raw spreadsheet data (Figure~\ref{fig:ux}a). Users can zoom into or out of the overview to obtain a \new{fine or coarse-}grained 
perspective of the data distribution. 
The distribution at each granularity 
is captured by a histogram,
enabling users to assimilate the data \new{via summary
statistics}. 
Each bin (group) of the histogram is 
mapped to a collection of rows in the spreadsheet. 
Cumbersome scrolling operations are eliminated in favor
of a few clicks on the overview interface. Instead of steering to analyze the data,
users can issue formulae
on the overview with interactions similar to pivot table construction,
and view results on a separate {\em aggregate column},
alongside the overview (Figure~\ref{fig:ux}b). 
In this manner,
users can issue formulae on different
subsets of the data while remaining on the same screen,
reducing visual discontinuity.
\noah ensures that there is coordination between
the overview and the spreadsheet:
for example, panning and zooming on the overview
are reflected on the spreadsheet by displaying
the spreadsheet data corresponding to the bin
currently in focus in the overview.
Finally, \noah automatically creates
contextual and historical information
(Figure~\ref{fig:ux}d and~\ref{fig:ux}e)
while displaying visual cues
(Figure~\ref{fig:ux}f and~\ref{fig:ux}g)
so that users don't lose context during navigation.

The primary contribution of our work is twofold:
\squishlist
\item We formalize the design of a general
navigation (overview+detail) interface
for exploration and analysis of large spreadsheets.
We realize this design in the form of \noah,
a plugin to an existing web-based spreadsheet tool,
ensuring that interactions supported by \noah
complement existing spreadsheet operations.
\item We conduct a user study to evaluate
the benefits and limitations of this plugin.
\newsaj{The study required users to
perform tasks that were representative of
popular spreadsheet operations,
using both Microsoft Excel and \noah.}
The study shows that compared to Microsoft Excel, participants were able to complete spreadsheet
tasks involving navigation 
correctly and quickly in \noah. Participants made {\bf $2.5\times$ fewer} mistakes while being {\bf $2\times$ faster} with a \noah-integrated spreadsheet than with Excel. \agp{mention how these tasks were representative or general}
\squishend

\toappendix{The rest of the paper is organized as follows: In Section~\ref{sec:usage} we first present a case study of \noah being applied to a real-world scenario.  
In Section~\ref{sec:related}, we discuss related work. We then explain the design considerations for developing \noah in Section~\ref{sec:design}. In Section~\ref{sec:ui}, we present \noah’s user interface and supported operations. We then discuss the design and outcomes of a user study evaluating the effectiveness of \noah in Section~\ref{sec:study}. Finally, we discuss the limitations of \noah and possible future enhancements in Section~\ref{sec:discuss}.}